Using Linear Regression to Examine the Correlation Between Access to Drinking Water and Life Expectancy

### 2022-12-09 Paper.tex
fixing citations

These results highlight the importance for access to clean drinking water, as we saw a positive correlation between access to safe water and life expectancy. We also saw some variation in the residuals, meaning that the model predicted a different outcome than the data points observed. The level of variation is a discrepancy between the predicted life expectancy of the model and the actual observed life expectancy from the data. This most likely has to do with the varying factors among different countries including other health factors such as air quality, sanitation, living conditions, and political stability. More conclusive research would have to be done to verify this. Possible methods such as multivariate regression and an analysis of covariance could be done. It is known that a person's health is affected by many factors, and a complete study to analyze just how much effect each factor has would be extensive. Though a study of this scale is improbable for this paper, such findings could revolutionize what public health organizations, governments, and citizens prioritize in relation to their health. This type of study has far-reaching effects and could not only improve a person's life expectancy but also their quality of life.

>>>>

Once the data was adjusted for correlating countries in the same year, the scatterplot suggested a positive linear relationship existed. Once the line of best fit was computed, this was confirmed. The slope of the least squares line is 0.279 with an intercept of 49.157. Theoretically, if $0\%$ of the population had access to clean drinking water, we would expect the average life expectancy at birth to be 49.157. Every one-unit increase in the percentage of access to drinking water results in a 0.279 year increase in the average life expectancy. These results highlight the importance for access to clean drinking water, as we saw a positive correlation between access to safe water and life expectancy.

\section{Limitations}

We saw some variation in the residuals, meaning that the model predicted a different outcome than the data points observed. The level of variation is a discrepancy between the predicted life expectancy of the model and the actual observed life expectancy from the data. This most likely has to do with the varying factors among different countries including other health factors such as air quality, sanitation, living conditions, and political stability. More conclusive research would have to be done to verify this. Possible methods such as multivariate regression and an analysis of covariance could be done. Moreover, this data does not take into account how far away the source of the safe drinking water is, if one is present. From the work done in \citet{2018waterandhealth}, "Across all countries in this study, results show that up to 40\% of the national population needs more than 30 minutes to fetch water irrespective of the type of water source used by the household...as reducing the collection burden required to fetch water is essential to enhancing water quantity available for households and improving general health and quality of life". It is known that a person's health is affected by many factors, and a complete study to analyze the effect of each factor would be extensive. Though a study of this scale is improbable for this paper, such findings could revolutionize what public health organizations, governments, and citizens prioritize in relation to their health. This type of study has far-reaching effects and could not only improve a person's life expectancy but also their quality of life.

\section{Conclusion}
